Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2023-02-06 manuscript.Rmd
typo

<!---many researcher do not subscribe to confirmatory agenda--->
We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
When researcher criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
It might be that psychology needs more confirmatory studies but preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.

@@

We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
<--AB: can you rewrite this and be more clear? Isn't this more about researchers who do not distinguish these concepts? -->
Subscribing to this distinction would mean to use preregistrations solely as a means to increase the theoretical risk researchers are willing to accept.
Undoubtedly, this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
<!--AB: maybe add "everything else being the same" or the specific conditions?-->
As we further argued, minimizing the uncertainty about theoretical risk (without necessarily increasing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.

>>>>

<!---many researcher do not subscribe to confirmatory agenda--->
We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
It might be that psychology needs more confirmatory studies but preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.

@@

We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
<!--AB: can you rewrite this and be more clear? Isn't this more about researchers who do not distinguish these concepts? -->
Subscribing to this distinction would mean to use preregistrations solely as a means to increase the theoretical risk researchers are willing to accept.
Undoubtedly, this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
<!--AB: maybe add "everything else being the same" or the specific conditions?-->
As we further argued, minimizing the uncertainty about theoretical risk (without necessarily increasing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
